### 2014-03-17 docs/computEL/liling.tex

\documentclass[11pt]{article}
\title{\textbf{universal corpus}}
\date{}
\begin{document}
\maketitle

\end{document}