### 2014-03-19 docs/computEL/liling.tex
added countings for wikipedia

\documentclass[11pt]{article}
\title{\textbf{universal corpus}}
\date{}
\begin{document}
\maketitle

\section{Corpus Coverage and Universality, something, something numbers...}

\begin{table}
\centering
    \begin{tabular}{l|cc}
    ~         & \#Languages & \#Families \\ \hline
    ODIN      & 1218       & 101       \\
    Omniglot  & 134        & 21        \\
    UDHR      & 355        & 47        \\
    Wikipedia & 209        & 22       \\
    \end{tabular}
\caption{Corpus Coverage}
\end{table}

According to ethnologue
Total Living languages: 7105
Total Living language families: 147

\end{document}