### 2014-03-20 docs/computEL/data.tex
Update data.tex

\section{The Data}\label{sec:data}

\subsection{Representation and Universality} \label{sec:stats}

According to Ethnologue, there are 7105 living languages, and 147 living language families. Across all our data sources, we manage to cover 1347 languages in 106 families, which represents 19.0\% of the world's languages. To get a better idea of the kinds of languages represented, we give a breakdown according to their EGIDS scores (Expanded Graded Intergenerational Disruption Scale) in figure \textcolor{blue}{[insert reference here when it's ready]}. The values in each cell have been colored according to proportion of languages represented, with green indicated good coverage and red poor. It's interesting to note that vigorous languages (6a) are poorly represented across all data sources, and worse than more endangered categories. In terms of language documentation, vigorous languages are less urgent goals than those in categories 6b and up, but this highlights an unexpected gap in linguistic resources.

%\begin{table}
%\centering
%    \begin{tabular}{l|cc}
%    ~         				& \#Languages & \#Families \\ \hline
%    ODIN      				& 1218       & 101       \\
%    Omniglot  				& 134        & 21        \\
%    UDHR      				& 355        & 47        \\
%    Wikipedia 				& 209        & 22       \\ \hline
%    \textbf{Combined}	& 1347			 & 106 
%    \end{tabular}
%\caption{Corpus Coverage}
%\end{table}

\begin{table}[h!]
\centering
    \begin{tabular}{l|rr|rr}
    ~         				& \#Languages & \#Families 	&\#tokens		& Size	\\ \hline
    ODIN      				& 1,217      & 101       		& 58,556		& 39 MB		\\
    Omniglot  				& 134        & 21        		&	6,749			& 677 KB	\\
    UDHR      				& 355        & 47        		&	33,117		& 5.2 MB	\\
    Wikipedia 				& 209        & 22       		&						& 37 GB		\\ \hline
    \textbf{Combined}	& 1,347			 & 106 
    \end{tabular}
\caption{Corpus Coverage}
\label{table:corpus}
\end{table}

\textcolor{blue}{[insert heatmap here]}

\subsection{Universal Corpus and Data Structure} \label{sec:structure}

Building on their previous paper, \newcite{abney2011data} describe the data structure they envisage for the universal corpus in more detail, distinguishing between \textbf{aligned texts} and \textbf{analysed texts}. Aligned texts consist of multiple parallel documents, aligned at the document, sentence, or word level. The collection of parallel documents as a whole is assigned a unique identifier, with each individual document labelled by this identifier and the corresponding language code.  A monolingual document is simply one for which the document ID only has a single language code associated with it.  Sentences and words within documents are also assigned identifiers (relative to the document), if alignment has been performed at that level.

Analyses texts, in addition to the raw text, contain more detailed annotations including parts of speech, morphological information, and syntactic relations.  This is stored as a table, where each row represents a word, and each column a type of information.  They propose the following columns: document ID, language code, sentence ID, word ID, wordform, lemma, morphological information, part of speech, gloss, head/governor, and relation/role.

Out of our data sources, three can be straightforwardly represented in their aligned text structure.  However, ODIN contains richer annotations, which are in fact difficult to fit into their proposal, which we discuss in section \ref{sec:odin} below.

\subsection{Data Sources} \label{sec:sources}

Although data size matters in general NLP, \emph{universality} is the utmost priority for a universal corpus. We chose to focus on the following data sources, because they include a large number of languages, include several parallel texts, and demonstrate a variety of data types which a linguist might encounter (structured, semi-structured, unstructured):

\begin{enumerate}
\item The Online Database of Interlinear Text (ODIN)
\item The Omniglot website
\item The Universal Declaration of Human Rights (UDHR)
\item Wikipedia
\end{enumerate}

Our resulting corpus runs the full gamut of text types outlined by Abney and Bird, ranging from single-language text (Wikipedia) to parallel text (UDHR and Omniglot) to IGTs (ODIN).  We describe each source in the following subsections.

\subsubsection{Universal Declaration of Human Rights}

The Universal Declaration of Human Rights (UDHR) is a document released by the United Nations in 1948, and represents the first global expression of human rights. It consists of 30 articles, amounting to about four pages of text. This is a useful document for CL, since it has been translated into a wide variety of languages, providing a highly parallel text.

\subsubsection{Wikipedia}

Wikipedia\footnote{url{http://www.wikipedia.org}} is a collaboratively-edited encyclopedia, appealing to use for NLP because of its large size and easy availability. At the time of writing, it contained 30.8 million articles in 287 languages, which provides a sizeable amount of clean monolingual text in a fairly wide range of languages. Text dumps are made regularly, and can be downloaded from \url{www.dumps.wikimedia.org}.

\subsubsection{Omniglot}

The Omniglot website\footnote{url{http://www.omniglot.com}} is an online encyclopedia of writing systems and languages. We chose to extract information from pages on \emph{`Useful foreign phrases'} and the \emph{`Tower of Babel'} story, both of which give us a parallel data in a reasonably large number of languages. The urls for these resources are:

\begin{itemize}[noitemsep]
\item \url{www.omniglot.com/language/phrases/*} 
\item \url{www.omniglot.com/babel/*}
\end{itemize}

\subsubsection{ODIN} \label{sec:odin}

ODIN (The Online Database of Interlinear Text) is a repository of interlinear glossed texts (IGTs) extracted from scholarly documents \cite{lewis2006odin,lewis2010odin}.  Compared to other resources, it is notable for the breadth of languages included, and the level of linguistic annotation.  An IGT consists of three lines: (i) the source, a sentence in a target language, (ii) the gloss, an analysis of each element in the source, and (iii) the translation, done at the sentence level. The gloss line can additionally include a number of linguistic terms, which means that the gloss can more properly described as being written in a metalanguage, rather than a natural language.  In ODIN, translations are into English, and glosses are written in an English-based metalanguage.  An accepted set of guidelines are given by the Leipzig Glossing Rules,\footnote{\url{http://www.eva.mpg.de/lingua/resources/glossing-rules.php}}, where morphemes within words are separated by hyphens (or equal signs, for clitics), and the same number of hyphens should appear in each word of the source and gloss.

In contrast, the proposed data structure for the universal corpus is aligned at the word level, and includes a specific list of relevant features which should be used to annotate words. When we try to adapt IGTs into this format, we run into certain problems.  Firstly, there is the problem that the most fundamental unit of analysis according to the Leipzig Glossing Rules is the morpheme, not the word.  Ideally, we should encode this information explicitly in a universal corpus, assigning a unique identifier to each morpheme (instead of, or in addition to each word). Indeed, \newcite{haspelmath2011segment} argues that there is no cross-linguistically valid definition of "word", which undermines the central position of words in the proposed data structure.

Secondly, it is unclear how to represent the gloss.  Since the gloss line is written in a natural language, we cannot treat it as a simple translation.  However, it is not straightforward to incorporate it into the proposed structure for analysed texts, either.  We could simply always map the gloss to the \textsc{gloss} field, but this would run against the suggestion that this field should "contain symbols from controlled vocabularies".  Indeed, this information might be more suited to the \textsc{morph, pos,} or \textsc{rel} fields, but it would be difficult to automatically process ODIN to determine which morphemes should be assigned to which field.

One possible resolution is to move all elements of the gloss written in capital letters to the \textsc{morph} field (as functional elements are usually annotated in this way), and all remaining elements to the \textsc{gloss} field.  However, this loses information, since we no longer no which morpheme has which meaning.  It would be ridiculous to throw away information in the IGTs, but to keep all information, we need to modify the \newcite{abney2011data}'s proposal.

The simplest solution we can see is to allow morphemes to be a level of structure in the universal corpus, just as documents, sentences, and words already are.  The overall architecture remains unchanged.  We must then decide how to represent the glosses.

Even though glosses in ODIN are based on English, having been extracted from English-language documents, this is not true of IGTs in general.  In particular, it is common for documentary linguists working on indigenous languages of the Americas to provide glosses and translations based on Spanish.  For this reason, we believe it would be wise to specific the language used to produce the gloss.  Since it is not quite the language itself, but a metalanguage, one solution would be to use new language codes that make it clear both that a metalanguage is being used, and also what natural language it is based on.  The five-letter code \texttt{gloss} cannot be confused with any code in any version of ISO 639 (which includes codes of length two to four).  Following the convention that subvarieties of a language are indicated with suffixes, we can append the code of the natural language being used.  For example, glosses into English and Spanish-based metalanguages would be given the codes \texttt{gloss-eng} and \texttt{gloss-spa}, respectively.

One benefit of this approach is that glossed texts are treated in exactly the same way as parallel texts.  There is a unique identifier for each morpheme, and glosses are stored under this identifier and the corresponding gloss code.  Furthermore, to motivate the important place of parallel texts in a universal corpus, Abney and Bird view translations into a high-resource reference language as a convenient surrogates of meaning. By the same reasoning, we can use glosses to provide a more detailed surrogate of meaning, only written in a metalanguage instead of a natural one.