### 2014-03-22 docs/computEL/paper.tex

\documentclass[11pt]{article}
\usepackage{acl2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

\bibliographystyle{acl}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\usepackage{enumitem}
\usepackage{listings}
\usepackage{fancyvrb}
\usepackage{color}
\definecolor{gray}{rgb}{0.4,0.4,0.4}
\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{orange}{rgb}{1,0.5,0}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  showstringspaces=false,
  commentstyle=\color{gray}\upshape
}

\lstdefinelanguage{XML}
{
  morestring=[b]",
  morestring=[s]{>}{<},
  morecomment=[s]{<?}{?>},
  stringstyle=\color{black},
  identifierstyle=\color{darkblue},
  keywordstyle=\color{cyan},
  morekeywords={xmlns,version,type}% list your attributes here
}

\newcommand{\bluenote}[1]{\textcolor{blue}{[#1]}}

\setlength\titlebox{5cm}

\title{SeedLing: Building and using a seed corpus \\ for the Human Language Project}
%Getting the Ball Rolling: Producing a Foundation Text for a Universal Corpus of the World's Languages}

% \author{Liling Tan, Guy Emerson, Susanne Fertmann, Alexis Palmer and Michaela Regneri \\
%   Universität des Saarlandes \\
%   Campus, 66123 Saarbrücken, Germany \\
%   {\tt liling.tan@uni-saarland.de, emerson@coli.uni-saarland.de,} \\
%   {\tt susfert@coli.uni-saarland.de}}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  A bread-coverage corpus such as the Human Language Project
  envisioned by \newcite{abney2010universal} would be a powerful
  resource for the study of endangered languages.
%, \bluenote{keep? from
 %   the perspective of both the documentary and the computational
  %  linguist}. 
Existing corpora are limited in the range of languages
  covered, in standardisation, or in machine-readability. In this
  paper we present SeedLing, a seed corpus for the Human Language
  Project. We first survey existing efforts to compile
  cross-linguistic resources, then describe our own approach. To build
  the foundation text for a universal corpus, we crawl and clean
  texts from several web sources that contain data from a large number
  of languages, and convert them into a standardised form consistent
  with the guidelines of \newcite{abney2011data}. The
  resulting corpus is more easily-accessible and machine-readable than
  any of the underlying data sources, and represents a significant
  base corpus for researchers to draw on and add to in the future. To show the utility of SeedLing for cross-lingual computational research, we use our data in the test application of clustering languages.
\end{abstract}

\input{intro}
\input{relwork}
%\input{liling}
\input{data}
\input{clean}
\input{cluster}
%\input{corpus}
\input{concl}

\bibliography{references}

\end{document}