### 2014-04-17 docs/computEL-final/cluster.tex

% \section{Detecting Similar Languages} \label{sec:cluster}

% When working on low-resource or endangered languages, computational and documentary linguists face the issue of lack of resources and knowledge about the language. Often, having knowledge about related or similar languages provides useful lexical, syntactic or morphological minimal pairs across languages and also helps in bootstrapping language models in NLP \cite{yarowsky:ngai:2001,xia2007multilingual}.

% Language classification can be carried out either: geneologically, mapping languages onto families trees depending on their historical ancenstry \cite{swadesh1952,starostin2010}; or typologically, grouping languages according to typological features \cite{georgi2010wals,daume2009}.

% To exemplify the use of the universal corpus for research on low-resource languages, we experimented with automatic detection of similar languages using hierarchical clustering with character ngrams and word unigrams features. Each language is represented by a vector of character bigrams and trigrams and word unigram features from the universal corpus. 
% \newline \newline
% \noindent \textbf{SHOW FANCY FIGURE OF CLUSTER}
% \newline \newline
% \textbf{Evaluate the cluster (quantitatively/qualitatively)}

%%%%%%%%%%%

\section{Detecting Similar Languages}

To exemplify the use of SeedLing for computational research that is relevant for low-resource languages, we experiment  with automatic detection of similar languages on the basis of our data. 
\label{sec:cluster} When working on low-resource or endangered 
languages, documentary and computational linguists alike face the
issue of lack of resources and knowledge about the language. Often,
having knowledge about related or similar languages provides useful
lexical, syntactic or morphological minimal pairs across
languages. What's more, information about relatedness between
languages can be useful for identifying data sources for high-resource
languages to be used in bootstrapping approaches such as those
described in \newcite{yarowsky:ngai:2001} or
\newcite{xia2007multilingual}.

Language classification can be carried out in a number of ways. Two common approaches are genealogical classification, mapping languages onto family trees according to their historical relatedness \cite{swadesh1952,starostin2010}; and typological classification, grouping languages according to certain linguistic features \cite{georgi2010wals,daume2009}. These approaches require linguistic analysis. By contrast, we use very simple features (character ngrams and word unigrams) extracted from SeedLing and an off-the-shelf hierarchical clustering algorithm.\footnote{\url{http://glaros.dtc.umn.edu/gkhome/views/cluto}} Specifically, each language is represented by a vector of frequencies of character bigrams, character trigrams, and word unigrams. Each of the three types of vector components is normalized by unit length.

% To exemplify the use of the universal corpus for research on low-resource languages, we experimented with automatic detection of similar languages using hierarchical clustering with character ngrams and word unigrams features. Each language is represented by a vector of frequencies of  character bigrams, character trigrams and word unigram features. Each of the three components in the vector is normalized by unit length.

% \begin{table*}[t]
% \begin{centering}

%     \begin{tabular}{l|ccc|ccc}
%     ~        & complete & ~       & ~       & ward    & ~       & ~       \\ \cline{2-7}
%     ~        & precision & recall       & f-score       & precision    & recall       & f-score      \\ \hline
%     distance & 0.0614   & 0.8565 & \textbf{0.1099} & 0.06140  & 0.8565 & \textbf{0.1099} \\
%     maxclust & 0.1925  & \textbf{0.0927} & 0.0692 & \textbf{0.1963} & 0.0905 & 0.0686  \\
%     \end{tabular}
% \caption{Comparing of clustering algorithms where number of clusters is set to genetic grouping on Ethnologue.}
% \end{centering}
% \label{tab:cluster}
% \end{table*}

\paragraph{Experimental Setup.}
We perform hierarchical/agglomerative clustering using a variety of linkage methods: (i) \texttt{single}, (ii) \texttt{complete} and (iii) \texttt{weighted}.
%(WPGMA, weighted pair group method with averaging). 
The \texttt{single} method calculates the distance between the newly formed clusters by assigning the minimal distance between the clusters and the \texttt{complete} method assigns the maximal distance between the newly formed clusters. The \texttt{weighted} method assigns the averaged distance between the newly formed cluster and its intermediate clusters. We set the number of clusters to 147, the number of top-level genetic groupings in Ethnologue (presumably the maximum number of language family clusters we would seek to induce).

% The \texttt{maxclust} criterion for flattening clusters from the hierarchy is achieved by finding the minimum threshold, \emph{r}, such that the cophenetic distance between any two original observations in the same flat cluster is no more than \emph{r} and not more than the number of clusters.

\begin{table}[t]
\begin{centering}

    \begin{tabular}{l|ccc}
    ~        & Precision & Recall       & F-score    \\ \hline
    \texttt{single} & 0.1958	& 0.6755	 & 0.1240  \\
	\texttt{complete} & \textbf{0.3153}	& 0.1699	 & \textbf{0.1420} \\
	\texttt{weighted} & 0.0614	& \textbf{0.8565}	 & 0.1099 \\ \hline
	\emph{random} & 0.1925 &	0.0927 & 	0.0692 \\
    \end{tabular}
\caption{Comparison of clustering algorithms} %where number of clusters is set to genetic grouping on Ethnologue.} 
\end{centering}
 \label{table:cluster}
\end{table}

\paragraph{Evaluation.}
There are many possible metrics to evaluate the quality of a clustering compared to a gold standard. 
\newcite{amigo2009metrics} propose a set of criteria which a clustering evaluation metric should satisfy, and demonstrate that most popular metrics fail to satisfy at least one of these criteria.  However, they prove that they are satisfied by the BCubed metric, which we adopt for this reason.  To calculate this, we find the induced cluster and gold standard class for each language, and calculate the F-score of the cluster compared to the class.  These F-scores are then averaged across all languages.

In table 2, we give results of our clustering experiments, comparing three clustering approaches to a random baseline.  The F-scores are comparable to those reported by \newcite{georgi2010wals}, even though we have only used surface features, while they used typological features taken from WALS.  This demonstrates that it possible for cross-linguistic research to be conducted even based on extremely shallow features.

% As well as performing clustering, we can view this as an information retrieval task: given a language, what are similar languages?  To do this, we found the languages with the closest vectors of n-grams and words.  Since BCubed is calculated averaging across languages and not clusters, we can use exactly the same calculation, using the set of nearby languages in place of a cluster.

% In table \ref{table:cluster}, we give results of our clustering experiments.  The F-scores are much higher than the random baseline, and comparable to the values reported by \newcite{georgi2010wals}, even though we have only used surface features, while they used typological features taken from WALS.  This demonstrates that it possible for cross-linguistic research to be conducted even based on extremely shallow features.

It is also worth noting that precision is higher than recall.  This is perhaps to be expected, given that related languages using wildly differing orthographies will appear to be very different.  Nonetheless, our system is reasonably capable of identifying those languages which are both related and also written similarly.