### 2014-05-19 docs/The Adventures of Sugali.tex

\documentclass[11pt]{article}

\title{\textbf{SuGaLi: A Quest to Identify the Languages of the World}}
\author{Susanne Fertmann, Guy Emerson, Liling Tan}
\date{}

\usepackage{hyperref}

\begin{document}

\maketitle

\section{Prologue}

\emph{``And they said, Go to, let us build us a corpus and a software, whose bytes may reach unto heaven; and let us assign a tag for each language, lest we be scattered abroad upon the encodings and data sparsity issues of the whole earth..."} - The Tower of Sugarlike. \\

\noindent It started with the ambition to create a software in the name of Natural Language Processing for Low Resource Language. Three brave scientists pit their wits against the world to show that they can create a software that can identify any language written in the mortal world.

Little did they know that the task perils of bytes that stood before them. Little had they realized the growth of knowledge and potential of what they had created. 
\\

\noindent

\newpage
\section{Bytes by Bytes: Building the Universal Corpus}

\emph{Ashes to ashes, dust to dust, \\
unlike my body, my corpus can't rust. \\
Bytes by bytes, the corpus swells, \\
crawling and cleaning, thinking all is well. \\}
- Songs of Corpora Minstrel \\

\noindent Desperate for data, there was none. The poor scientists had to synthesize one. Searching high and low, the scientists surfed, being extra careful not stepping on copyrights turf.

Crawl and clean, the scientists gleaned. Sifting the finest bytes, they toiled through the night. And atlas, a \texttt{SeedLing} corpus, the scientists plants.

\section{The Phantom Menace: Encodings}
\emph{``Behold, the encoding is one, and they have all one unicode; and this they begin to do: and now nothing will be restrained from them, which they have imagined to do. Go to, let us go down, and there confound their encodings, that they may not understand one another's text."} - The Tower of Sugarlike. \\

\noindent Plague by the miasma of encodings from the different files. Standardizing them wasn't as easy as lego tiles.

\section{The Menace Phantom: Language Codes}

\emph{``Behold, the language code is one, and they have all one ISO; and this they begin to do: and now nothing will be restrained from them, which they have imagined to do. Go to, let us go down, and there confound their ISOs, that they may not differentiate one another's text."} - The Tower of Sugarlike. \\

\
\\
\
\\

\bibliography{literature}{}
\bibliographystyle{plain}

\end{document}